---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# openalexR

<!-- badges: start -->
<!-- badges: end -->

The goal of openalexR is to gather bibliographic metadata about publications, authors, venues, institutions and concepts from OpenAlex using API.

OpenAlex is a fully open catalog of the global research system. 
It's named after the ancient [Library of Alexandria](https://en.wikipedia.org/wiki/Library_of_Alexandria).
The OpenAlex dataset describes scholarly entities and how those entities are connected to each other.
There are five types of entities:

* **Works** are papers, books, datasets, etc; they cite other works

* **Authors** are people who create works

* **Venues** are journals and repositories that host works

* **Institutions** are universities and other orgs that are affiliated with works (via authors)

* **Concepts** *tag* Works with a topic

(source:  [OpenAlex website](https://openalex.org))

&nbsp; 

## Installation

You can install the developer version of the openalexR from [GitHub](https://github.com) with:

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("massimoaria/openalexR")
```

You can install the released version of openalexR from [CRAN](https://CRAN.R-project.org) with:

```{r eval=FALSE}
install.packages("openalexR")
```

## Package loading

```{r}
library(openalexR)
```

# openalexR overview

The basic idea of openalexR is to provide three main functions helping the user to:

* Create a query by passing one or more arguments to a function (function **oaQueryBuild**)

* Gather a collection of entities in JSON format (function **oaApiRequest**)

* Transform the JSON in a data frame (similar to an excel sheet) can be used as input in a bibliometric or science mapping analysis (e.g. using the bibliometrix package) (function **oa2df**)

&nbsp;

OpenAlex defined a custom query language based on entity type. You can choose to write a valid query using that language or, in alternative, using the function **oaQueryBuild**.

**oaQueryBuild** generates a valid query, written following the OpenAlex API language, from a set of arguments provided by the user.

The function **oaApiRequest** downloads a collection of entities matching the query created by **oaQueryBuild** or manually written by the user. The function will return a JSON object in a list format.

Finally, the function **oa2df** converts  the JSON object in classical bibliographic data frame.

&nbsp; 

## Get full records through entity IDs.

&nbsp; 

### Query to obtain all information about a single publications

The following paper:

    Aria, M., & Cuccurullo, C. (2017). bibliometrix: 
    An R-tool for comprehensive science mapping analysis. 
    Journal of informetrics, 11(4), 959-975.

is associated to the OpenAlex-id **W2755950973**.

In this example, we need to pass a single argument to the function, that is, the identifier of the entity to download: identifier = "W2755950973".

```{r}
query_work <- oaQueryBuild(
  identifier = "W2755950973",
  entity = "works"
)

cat(query_work)
```


As results, **oaQueryBuild** returns the query string including the OpenAlex endpoint API server address. You should change it by using the argument "endpoint = *address*"

The function **oaApiRequest** downloads the bibliographic records matching the query.


```{r}
res <- oaApiRequest(
  query_url = query_work
)

df <- oa2df(res, entity = "works")

dplyr::glimpse(df)
```

&nbsp; 

### Query to obtain all information about a single publications using external id formats

OpenAlex endpoint accepts an OpenAlex ID, but many external IDs (*e.g.*, DOI, ISSN) are accepted as well, in several formats.


**DOI (Digital Object Identifier)**

We can get a publication record through its DOI using the format **doi:***doi identifier*. 
Example:

```{r}
query_work <- oaQueryBuild(
  identifier = "doi:10.1016/j.joi.2017.08.007",
  entity = "works"
)

cat(query_work)
```

```{r}
res <- oaApiRequest(
  query_url = query_work
)
df <- oa2df(res, entity = "works")

dplyr::glimpse(df)
```



**Persistent Identifiers (PIDs)**

Many persistent identifiers (PIDs) are canonically expressed as a URL that will take you to the thing being identified. Where these URL formats exist, OpenAlex treats them as the canonical ID, and also accepts them as valid IDs.
Example:

```{r}
query_work <- oaQueryBuild(
  identifier = "doi:https://doi.org/10.1016/j.joi.2017.08.007",
  entity = "works"
)

cat(query_work)

res <- oaApiRequest(
  query_url = query_work
)

df <- oa2df(res, entity = "works")

dplyr::glimpse(df)
```

### Query to obtain all information about a two o more publications

To download the records of two o more identifiers through a single query, we can recursively apply **oaApiRequest** to each id using the function **lapply**.

```{r}
ids <- c("W2755950973", "W3005144120")

res <- lapply(ids, function(x) {
  oaApiRequest(
    query_url = oaQueryBuild(
      identifier = x,
      entity = "works"
    )
  )
})

df <- oa2df(res, entity = "works")

dplyr::glimpse(df)
```

### Query to obtain all information about a single author

The author Massimo Aria is associated to the OpenAlex-id A923435168.

```{r}
query_author <- oaQueryBuild(
  identifier = "A923435168",
  entity = "authors"
)

cat(query_author)
```

```{r}
res_author <- oaApiRequest(
  query_url = query_author,
  total.count = FALSE,
  verbose = FALSE
)
```

```{r}
df <- oa2df(res_author, entity = "authors")

dplyr::glimpse(df)
```

&nbsp;

&nbsp;

## Get all entities matching a set of inclusion/exclusion criteria (filters) 

In most cases, we are interested in downloading a collection of items that meet one or more inclusion/exclusion criteria (filters).

&nbsp;

In this case, the query definition will not be based on a single identifier but the choice of the entity type (usually "works") and one or more filters about this entity.

Filters narrow the list down to just entities that meet a particular condition--specifically, a particular value for a particular attribute. Supported attributes for each endpoints are listed on [OpenAlex API Documentation Website](https://docs.openalex.org/api/get-lists-of-entities#filter).

Filters are formatted thusly: **attribute:***value*. You set them using the *?filter* query parameter. Filters are case-insensitive.

&nbsp;

Each endpoint supports its own list of filters. Here they are, by endpoint:

&nbsp;

**/works filters**

You can filter using these attributes of the Works object. 

* display_name.search (alias: title.search)

* publication_year

* publication_date

* from_publication_date

* to_publication_date

* host_venue.issn

* authorships.author.id (alias: author.id)

* type

etc.

You can find more documentation about each attribute on the [OA Documentation Work page](https://docs.openalex.org/about-the-data/work).

&nbsp;

**/authors filters**

You can filter using these attributes of the Authors object. 

* display_name.search

* works_count

* cited_by_count

* last_known_institution.id

etc.

You can find more documentation about each attribute on the [OA Documentation Author page](https://docs.openalex.org/about-the-data/author).

&nbsp;

**/venues filters**

You can filter using these attributes of the Venue object. 

* display_name.search

* issn

* works_count

* cited_by_count

etc. 

You can find more documentation about each attribute on the [OA Documentation Venue page](https://docs.openalex.org/about-the-data/venue).

&nbsp;

**/institutions filters**

You can filter using these attributes of the Institution object. 

* display_name.search

* country_code

* type

* works_count
 
* cited_by_count

* x_concepts.id

You can find more documentation about each attribute on the  [OA Documentation Institution page](https://docs.openalex.org/about-the-data/institution).

&nbsp;

**/concepts filters**

You can filter using these attributes of the Concept object. You can find more documentation about each attribute on the Concept page.

* display_name.search

* level

* works_count

* cited_by_count

* ancestors.id

You can find more documentation about each attribute on the [OA Documentation Concept page](https://docs.openalex.org/about-the-data/concept).

&nbsp;

Below we show some examples of filters in use.

&nbsp;

### Filters based on string matching

**We want to download all works, cited more than 50 times, published between 2020 and 2021, which include the strings "bibliometric analysis" or "science mapping" in the title.**

To do that, we have to set filters about three attributes: title content (*"title.search"*), starting date for publication (*"from_publication_date"*), and ending date for publication (*"to_publication_date"*).

Starting and ending dates can be passed to the function **oaQueryBuild** using the arguments *date_from* and *date_to*. The format is YYYY-MM-DD.

The other attributes can be passed to the function through the argument *filter*.  

When an attribute has more than one item, these can be separated by the boolean operator OR represented by the symbol **|** .

On the contrary, different attributes have to be separated by commas.

e.g. **filter = 'title.search:"bibliometric analysis"|"science mapping", cited_by_count:>50'**

where:

* 'title.search:"bibliometric analysis"|"science mapping'

means all works containing the string "bibliometric analysis" OR "science mapping" in the publication title.

and:

* cited_by_count:>50

means all works cited more than 10 times.

The whole filter **'title.search:"bibliometric analysis"|"science mapping",cited_by_count:>50'** 

can be read as:

    *"all works containing the string "bibliometric analysis" OR "science mapping" 
    in the publication title AND cited more than 50 times"*.



```{r}
query <- oaQueryBuild(
  identifier = NULL,
  entity = "works",
  filter = 'title.search:"bibliometric analysis"|"science mapping",cited_by_count:>50',
  date_from = "2020-01-01",
  date_to = "2021-12-31",
  search = NULL,
  # sort = "cited_by_count:desc",
  endpoint = "https://api.openalex.org/"
)
```


The **sort** argument indicates how results have to be sorted. 

In this example results are sorted by total citations in a descending order.

Setting the argument total.count=TRUE, the function **oaApiRequest** returns the number of items matching the query without downloading the collection. 

```{r}
res <- oaApiRequest(
  query_url = query,
  total.count = TRUE,
  verbose = FALSE
)

res$count
```

Then, we can download the collection:

```{r}
res <- oaApiRequest(
  query_url = query,
  total.count = FALSE,
  verbose = FALSE
)

## OpenAlex downloading [======================] 100% eta:  0s
```

and transform it into a data frame:

```{r}
df <- oa2df(res, entity = "works")
dplyr::glimpse(df)
```

&nbsp;

### Get all works citing a particular work.

We can download all publications citing another publication by using the filter attribute **cites:***id*.

For example, if we want to download all publications citing the article Aria and Cuccurullo (2017), we have just to set the argument filter as:

**filter = "cites:W2755950973"**

where *"W2755950973"* is the OA id for the article by Aria and Cuccurullo.

```{r}
query1 <- oaQueryBuild(
  identifier = NULL,
  entity = "works",
  filter = "cites:W2755950973",
  date_from = NULL,
  date_to = NULL,
  # sort = NULL
)

res1 <- oaApiRequest(
  query_url = query1,
  total.count = TRUE,
  verbose = FALSE
)
```

This query will return a collection of `r res1$count` publications.
Let's to download it:

```{r}
res1 <- oaApiRequest(
  query_url = query1,
  total.count = FALSE,
  verbose = FALSE
)

# OpenAlex downloading [=====================] 100% eta:  0s
```

And then to convert it into a data frame:

```{r}
df <- oa2df(res, entity = "works")

dplyr::glimpse(df)
```

&nbsp;

### Get all authors matching a set of filters

We want download all authors' records of scholars who work at the University of Naples Federico II (OpenAlex ID: I71267560) and who have published more than 499 works:

```{r}
query_author <- oaQueryBuild(
  identifier = NULL,
  entity = "authors",
  filter = "last_known_institution.id:I71267560,works_count:>499"
)
```

Check how many records match the query:

```{r}
res <- oaApiRequest(
  query_url = query_author,
  total.count = TRUE,
  verbose = FALSE
)
res$count

## 34
```

Then, we download and convert the collection:

```{r}
res <- oaApiRequest(
  query_url = query_author,
  total.count = FALSE,
  verbose = FALSE
)
df <- oa2df(res, entity = "authors")

dplyr::glimpse(df)

```

&nbsp;

### Get all institutions matching a set of filters

We want download all records regarding Italian institutions (country_code:it) that are classified as educational (type:education):

```{r}
query_inst <- oaQueryBuild(
  entity = "institutions",
  filter = "country_code:it,type:education"
)
```

We check how many records match the query:

```{r}
res <- oaApiRequest(
  query_url = query_inst,
  total.count = TRUE,
  verbose = TRUE
)
res$count
```

Then we download and convert the collection:

```{r}
res <- oaApiRequest(
  query_url = query_inst,
  total.count = FALSE,
  verbose = TRUE
)
df <- oa2df(res, entity = "institutions")

dplyr::glimpse(df)
```

&nbsp;

### Get all venues matching a set of filters

We want download all records regarding journals that have published more than 100,000 works:

```{r}
query_venue <- oaQueryBuild(
  entity = "venues",
  filter = "works_count:>100000"
)
```

We check how many records match the query:

```{r}
res <- oaApiRequest(
  query_url = query_venue,
  total.count = TRUE,
  verbose = TRUE
)
res$count
```

Then we download and convert the collection:

```{r}
res <- oaApiRequest(
  query_url = query_venue,
  total.count = FALSE,
  verbose = TRUE
)

df <- oa2df(res, entity = "venues")

dplyr::glimpse(df)

```

&nbsp;

### Get all concepts matching a set of filters

We want to download the records of all the concepts that concern at least one million works:

```{r}
query_concept <- oaQueryBuild(
  entity = "concepts",
  filter = "works_count:>1000000"
)
```

We check how many records match the query:

```{r}
res <- oaApiRequest(
  query_url = query_concept,
  total.count = TRUE,
  verbose = TRUE
)
res$count

## 112
```

Then we download and convert the collection:

```{r}
res <- oaApiRequest(
  query_url = query_concept,
  total.count = FALSE,
  verbose = TRUE
)

df <- oa2df(res, entity = "concepts")

dplyr::glimpse(df)
```

## Convert an OpenAlex data frame to a bibliometrix object

The bibliometrix R-package (https://www.bibliometrix.org) provides a set of tools for quantitative research in bibliometrics and scientometrics. It is written in the R language, which is an open-source environment and ecosystem. 

Today it represents one of the most used science mapping software in the world. In a recent survey on bibliometric analysis tools, Moral-Muñoz et al. (2020) wrote: “At this moment, maybe Bibliometrix and its Shiny platform contain the more extensive set of techniques implemented, and together with the easiness of its interface, could be a great software for practitioners”.

The function **oa2bibliometrix** converts a bibliographic data frame of works into a bibliometrix object. This object can be used as input collection of a science mapping workflow.


```{r}
query1 <- oaQueryBuild(
  identifier = NULL,
  entity = "works",
  filter = "cites:W2755950973",
  date_from = "2022-01-01",
  date_to = "2022-03-31",
  # sort = NULL
)

res1 <- oaApiRequest(
  query_url = query1,
  total.count = TRUE,
  verbose = FALSE
)
```

This query will return a collection of `r res1$count` publications.
Let's download it:

```{r}
res1 <- oaApiRequest(
  query_url = query1,
  total.count = FALSE,
  verbose = FALSE
)
```

Convert it into a data frame and then into a bibliometrix object:

```{r}
df <- oa2df(res1, entity = "works")
M <- oa2bibliometrix(df)
dplyr::glimpse(M)
```
