---
title: "A Brief Introduction to openalexR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A Brief Introduction to openalexR}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

 

[*https://github.com/ropensci/openalexR*](https://github.com/ropensci/openalexR){.uri}

*Latest version: `r packageVersion("openalexR")`, `r Sys.Date()`*

 

**by Massimo Aria**

Full Professor in Social Statistics

PhD in Computational Statistics

Laboratory and Research Group STAD Statistics, Technology, Data Analysis

Department of Economics and Statistics

University of Naples Federico II

email [aria\@unina.it](mailto:aria@unina.it){.email}

<https://massimoaria.com>

 

# An R-package to gather bibliographic data from OpenAlex

**openalexR** helps you interface with the [OpenAlex](https://openalex.org) API to retrieve bibliographic infomation about publications, authors, institutions, sources, funders, publishers, topics and concepts with 5 main functions:

-   `oa_query()`: generates a valid query, written following the OpenAlex API syntax, from a set of arguments provided by the user.

-   `oa_request()`: downloads a collection of entities matching the query created by `oa_query()` or manually written by the user, and returns a JSON object in a list format.

-   `oa2df()`: converts the JSON object in classical bibliographic tibble/data frame.

-   `oa_fetch()`: composes three functions above so the user can execute everything in one step, *i.e.*, `oa_query |> oa_request |> oa2df`

-   `oa_random()`: to get random entity, e.g., `oa_random("works")` gives a different work each time you run it

```{r warning=FALSE, message=FALSE}
library(openalexR)
library(dplyr)
```

## Works (think papers, publications)

This paper:

```         
Aria, M., & Cuccurullo, C. (2017). bibliometrix: 
An R-tool for comprehensive science mapping analysis. 
Journal of informetrics, 11(4), 959-975.
```

is associated to the [OpenAlex-id](https://docs.openalex.org/api-entities/works/work-object#id) **W2755950973**. If you know your paper's OpenAlex ID, all you need to do is passing `identifier = <openalex id>` as an argument in `oa_fetch()`:

```{r}
paper_id <- oa_fetch(
  identifier = "W2755950973",
  entity = "works",
  verbose = TRUE
)
dplyr::glimpse(paper_id)
```

`oa_fetch()` is a composition of functions: `oa_query |> oa_request |> oa2df`. As results, `oa_query()` returns the query string including the OpenAlex endpoint API server address (default). `oa_request()` downloads the bibliographic records matching the query. Finally, `oa2df()` converts the final result list to a tibble. The final result is a complicated tibble, but we can use `show_works()` to display a simplified version:

```{r}
paper_id %>% 
  show_works() %>%
  knitr::kable()
```

### External id formats

OpenAlex endpoint accepts OpenAlex IDs and other [external IDs](https://docs.openalex.org/how-to-use-the-api/get-single-entities#id-formats) (*e.g.*, DOI, ISSN) in several formats, including Digital Object Identifier (DOI) and Persistent Identifiers (PIDs).

```{r}
oa_fetch(
  # identifier = "https://doi.org/10.1016/j.joi.2017.08.007", # would also work (PIDs)
  identifier = "doi:10.1016/j.joi.2017.08.007",
  entity = "works"
) %>% 
  show_works() %>%
  knitr::kable()
```

### More than one publications/authors

[https://api.openalex.org/authors/https://orcid.org/](https://api.openalex.org/authors/https://orcid.org/0000-0003-1613-5981)

If you know the OpenAlex IDs of these entities, you can also feed them into the `identifier` argument.

```{r}
oa_fetch(
  identifier = c("W2741809807", "W2755950973"),
  # identifier = c("https://doi.org/10.1016/j.joi.2017.08.007", "https://doi.org/10.1016/j.joi.2017.08.007"), # TODO
  entity = "works",
  verbose = TRUE
) %>% 
  show_works() %>%
  knitr::kable()

```

However, if you only know their external identifies, say, DOIs, you would need to use `doi` as a filter (either the canonical form with <https://doi.org/> or without should work):

```{r}
oa_fetch(
  # identifier = c("W2741809807", "W2755950973"),
  doi = c("10.1016/j.joi.2017.08.007", "https://doi.org/10.1093/bioinformatics/btab727"),
  entity = "works",
  verbose = TRUE
) %>% 
  show_works() %>%
  knitr::kable()
```

### Filters

In most cases, we are interested in downloading a collection of items that meet one or more inclusion/exclusion criteria (filters). Supported filters for each entity are listed [here](https://ropensci.github.io/openalexR/articles/Filters.html).

**Example**: We want to download all works published by a set of authors. We can do this by filtering on the authorships.author.id/author.id or authorships.author.orcid/author.orcid attribute (see more on [works attributes](https://ropensci.github.io/openalexR/articles/Filters.html#works)):

```{r}
oa_fetch(
  entity = "works",
  author.id = c("A5048491430", "A5023888391"),
  verbose = TRUE
) %>% 
  show_works() %>% 
  knitr::kable()

orcids <- c("0000-0003-3737-6565", "0000-0002-8517-9411")
canonical_orcids <- paste0("https://orcid.org/", orcids)
oa_fetch(
  entity = "works",
  author.orcid = canonical_orcids,
  verbose = TRUE
) %>% 
  show_works() %>% 
  knitr::kable()
```

**Example**: We want to download all works that have been cited more than 50 times, published between 2020 and 2021, and include the strings "bibliometric analysis" or "science mapping" in the title. Maybe we also want the results to be sorted by total citations in a descending order.

Setting the argument `count_only = TRUE`, the function `oa_request()` returns the number of items matching the query without downloading the collection.

```{r}
oa_fetch(
  entity = "works",
  title.search = c("bibliometric analysis", "science mapping"),
  cited_by_count = ">50", 
  from_publication_date = "2020-01-01",
  to_publication_date = "2021-12-31",
  options = list(sort = "cited_by_count:desc"),
  count_only = TRUE,
  verbose = TRUE
)
```

We can now download the records and transform it into a tibble/data frame by setting `count_only = FALSE` (also the default value):

```{r}
oa_fetch(
  entity = "works",
  title.search = c("bibliometric analysis", "science mapping"),
  cited_by_count = ">50", 
  from_publication_date = "2020-01-01",
  to_publication_date = "2021-12-31",
  options = list(sort = "cited_by_count:desc"),
  count_only = FALSE
) %>%
  show_works() %>%
  knitr::kable()
```

Read on to see how we can shorten these two function calls.

## Authors

Similarly to work, we can use identifier to pass in authors' OpenAlex ID.

**Example**: We want more information on authors with IDs A5069892096 and [A5023888391](https://openalex.org/A5023888391).

```{r}
oa_fetch(
  identifier = c("A5069892096", "A5023888391"),
  verbose = TRUE
) %>%
  show_authors() %>%
  knitr::kable()
```

**Example**: We want download all authors' records of scholars who work at the University of Naples Federico II (OpenAlex ID: I71267560) and who have published more than 499 works.

Let's first check how many records match the query, then set `count_only = FALSE` to download the entire collection. We can do this by first defining a list of arguments, then adding `count_only` (default `FALSE`) to this list:

```{r}
my_arguments <- list(
  entity = "authors",
  last_known_institutions.id = "I71267560",
  works_count = ">499"
  )

do.call(oa_fetch, c(my_arguments, list(count_only = TRUE)))
do.call(oa_fetch, my_arguments) %>% 
  show_authors() %>%
  knitr::kable()
```

You can also use other filters such as `display_name`, `has_orcid`, and `orcid`:

```{r}
oa_fetch(
  entity = "authors",
  display_name.search = "Massimo Aria",
  has_orcid = "true"
) %>%
  show_authors() %>%
  knitr::kable()

oa_fetch(
  entity = "authors",
  orcid = "0000-0002-8517-9411"
) %>%
  show_authors() %>%
  knitr::kable()
```

## Institutions

**Example**: We want download all records regarding Italian institutions (country_code:it) that are classified as educational (type:education). Again, we check how many records match the query then download the collection:

```{r}
italian_insts <- list(
  entity = "institutions",
  country_code = "it",
  type = "education",
  verbose = TRUE
)

do.call(oa_fetch, c(italian_insts, list(count_only = TRUE)))
dplyr::glimpse(do.call(oa_fetch, italian_insts))
```

## Concepts (think theme, keywords)

**Example**: We want to download the records of all the concepts that concern at least one million works:

```{r}
popular_concepts <- list(
  entity = "concepts",
  works_count = ">1000000",
  verbose = TRUE
)

do.call(oa_fetch, c(popular_concepts, list(count_only = TRUE)))
dplyr::glimpse(do.call(oa_fetch, popular_concepts))
```

## Other examples

*Get all works citing a particular work*

We can download all publications citing another publication by using the filter attribute **cites**.

For example, if we want to download all publications citing the article Aria and Cuccurullo (2017), we have just to set the argument filter as `cites = "W2755950973"` where "W2755950973" is the OA id for the article by Aria and Cuccurullo.

```{r}
aria_count <- oa_fetch(
  entity = "works",
  cites = "W2755950973",
  count_only = TRUE,
  verbose = TRUE
) 
aria_count
```

This query will return a collection of `r aria_count["count"]` publications. Among these articles, let's download the ones published in the following year:

```{r}
oa_fetch(
  entity = "works",
  cites = "W2755950973",
  publication_year = 2018,
  count_only = FALSE,
  verbose = TRUE
) %>% 
  dplyr::glimpse()
```

## Convert an OpenAlex data frame to a bibliometrix object

The bibliometrix R-package (<https://www.bibliometrix.org>) provides a set of tools for quantitative research in bibliometrics and scientometrics. Today it represents one of the most used science mapping software in the world. In a recent survey on bibliometric analysis tools, Moral-Muñoz et al. (2020) wrote: "At this moment, maybe Bibliometrix and its Shiny platform contain the more extensive set of techniques implemented, and together with the easiness of its interface, could be a great software for practitioners".

The function **oa2bibliometrix** converts a bibliographic data frame of works into a bibliometrix object. This object can be used as input collection of a science mapping workflow.

```{r}
bib_ls <- list(
  identifier = NULL,
  entity = "works",
  cites = "W2755950973",
  from_publication_date = "2022-01-01",
  to_publication_date = "2022-03-31"
)

do.call(oa_fetch, c(bib_ls, list(count_only = TRUE)))
do.call(oa_fetch, bib_ls) %>% 
  oa2bibliometrix() %>% 
  dplyr::glimpse()
```

# About OpenAlex

[OpenAlex](https://openalex.org) is a fully open catalog of the global research system. It's named after the ancient [Library of Alexandria](https://en.wikipedia.org/wiki/Library_of_Alexandria). The OpenAlex dataset describes scholarly entities and how those entities are connected to each other. There are five types of entities:

-   **Works** are papers, books, datasets, etc; they cite other works

-   **Authors** are people who create works

-   **Institutions** are universities and other orgs that are affiliated with works (via authors)

-   **Concepts** *tag* Works with a topic

## Acknowledgements

Package hex was made with [Midjourney](https://www.midjourney.com/home/) and thus inherits a [CC BY-NC 4.0 license](https://creativecommons.org/licenses/by-nc/4.0/legalcode).
