---
title: "About the tibble output"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(openalexR)
library(dplyr)
```

The default output from an `oa_fetch` call is a tibble.
This object type allows each row to be one unit of entity (article, institution, etc.), which is often helpful for downstream wrangling.
It combines complex output elements in [list columns](https://dcl-prog.stanford.edu/list-columns.html), which can be **extracted** or **exploded** with `dplyr::rowwise` or `purrr::map`.

Of course, you can instead set `output = "list"` in your `oa_fetch` call if you are more comfortable working with list.

## Example 1: institutions

Suppose we queried Open Alex to obtain information on large Canadian institutions and now want to **extract their latitudes and longitudes**.

```{r}
institutions <- oa_fetch(
  entity = "institutions",
  country_code = "CA",
  cited_by_count = ">4000000",
  verbose = TRUE,
  count_only = FALSE
)

head(institutions)
```

We present three different approaches below.

### dplyr::rowwise

The use of [`rowwise`](https://dplyr.tidyverse.org/articles/rowwise.html) used to be [discouraged](https://community.rstudio.com/t/dplyr-alternatives-to-rowwise/8071), but the tidyverse team has now recognized its importance in many workflows, and so `rowwise` is here to stay.
We think `rowwise` pairs very naturally with our list columns output.

```{r}
institutions %>%
  rowwise() %>%
  mutate(
    name = display_name,
    openalex_id = stringr::str_extract(id, "I\\d+"),
    lat = geo$latitude,
    lon = geo$longitude,
    .keep = "none"
  )
```

### purrr::map

Alternatively, you can use any function in the `purrr::map` family.
As you can see below, the syntax is a little less natural, but you *may* gain some performance advantage if you have a really large dataframe.

```{r}
institutions %>%
  mutate(
    name = display_name,
    openalex_id = stringr::str_extract(id, "I\\d+"),
    lat = purrr::map_dbl(geo, ~ .x$latitude),
    lon = purrr::map_dbl(geo, ~ .x$longitude),
    .keep = "none"
  )
```

### base::lapply

Similar to the purrr approach, you can use the base functions in the `lapply` family, for example:

```{r}
institutions %>%
  mutate(
    name = display_name,
    openalex_id = stringr::str_extract(id, "I\\d+"),
    lat = vapply(geo, function(x) x$latitude, numeric(1)),
    lon = vapply(geo, function(x) x$longitude, numeric(1)),
    .keep = "none"
  )

```

## Example 2: works

Suppose we have a tibble of **works** output and would like to find the institutions corresponding with the works' authors.
In this case, each work may have more than one affiliated institutions.
We will store the result in a list column:

```{r}
works <- oa_fetch(
  entity = "works",
  title.search = c("bibliometric analysis", "science mapping"),
  cited_by_count = ">100", 
  from_publication_date = "2020-01-01",
  to_publication_date = "2021-01-31",
  sort = "cited_by_count:desc",
  count_only = FALSE
)

multi_insts <- works %>% 
  rowwise() %>% 
  mutate(
    openalex_id = stringr::str_extract(id, "W\\d+"),
    institutions = list(unique(author$institution_display_name)),
    .keep = "none"
  )

multi_insts

# institutions of the first work
multi_insts[1, "institutions"][[1]]

```
