---
title: "About the tibble output"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(openalexR)
library(dplyr)
```

The default output from an `oa_fetch` call is a tibble.
This object type allows each row to be one unit of entity (article, institution, etc.), which is often helpful for downstream wrangling.
It **simplifies** and combines complex output elements in [list columns](https://dcl-prog.stanford.edu/list-columns.html), which can be **extracted** or **exploded** with `dplyr::rowwise` or `purrr::map`.

**IMPORTANT**: in "flattening" the list output to a tibble, we have made a few decisions to simplify the original nested list such as retaining only one affiliation for each author if you're querying *Works* (details in [`oa_df` source code](https://github.com/massimoaria/openalexR/blob/main/R/oa2df.R)).
While the tibble output is sufficient in most use cases, you may need to obtain the original nested list with all the information on an entity for your special research problem.
If so, please specify `output = "list"` in your `oa_fetch` call.

## Example 1: institutions

Suppose we queried Open Alex to obtain information on large Canadian institutions and now want to **extract their latitudes and longitudes**.

```{r}
institutions <- oa_fetch(
  entity = "institutions",
  country_code = "CA",
  cited_by_count = ">4000000",
  verbose = TRUE,
  count_only = FALSE
)

head(institutions)
```

We present three different approaches below.

### dplyr::rowwise

The use of [`rowwise`](https://dplyr.tidyverse.org/articles/rowwise.html) used to be [discouraged](https://community.rstudio.com/t/dplyr-alternatives-to-rowwise/8071), but the tidyverse team has now recognized its importance in many workflows, and so `rowwise` is here to stay.
We think `rowwise` pairs very naturally with our list columns output.

```{r}
institutions %>%
  rowwise() %>%
  mutate(
    name = display_name,
    openalex_id = stringr::str_extract(id, "I\\d+"),
    lat = geo$latitude,
    lon = geo$longitude,
    .keep = "none"
  )
```

### purrr::map

Alternatively, you can use any function in the `purrr::map` family.
As you can see below, the syntax is a little less natural, but you *may* gain some performance advantage if you have a really large dataframe.

```{r}
institutions %>%
  mutate(
    name = display_name,
    openalex_id = stringr::str_extract(id, "I\\d+"),
    lat = purrr::map_dbl(geo, ~ .x$latitude),
    lon = purrr::map_dbl(geo, ~ .x$longitude),
    .keep = "none"
  )
```

### base::lapply

Similar to the purrr approach, you can use the base functions in the `lapply` family, for example:

```{r}
institutions %>%
  mutate(
    name = display_name,
    openalex_id = stringr::str_extract(id, "I\\d+"),
    lat = vapply(geo, function(x) x$latitude, numeric(1)),
    lon = vapply(geo, function(x) x$longitude, numeric(1)),
    .keep = "none"
  )
```

## Example 2: works

Suppose we have a tibble of **works** output and would like to find the institutions corresponding with the works' authors.
In this case, each work may have more than one affiliated institutions.

### Tibble output

Assuming that each author is affiliated with only one institution, we can call `oa_fetch` with the default `output = "tibble"`:

```{r}
works <- oa_fetch(
  entity = "works",
  title.search = c("bibliometric analysis", "science mapping"),
  cited_by_count = ">100",
  from_publication_date = "2020-01-01",
  to_publication_date = "2021-01-31",
  sort = "cited_by_count:desc",
  count_only = FALSE
)
```

We will store the result in a list column:

```{r}
multi_insts <- works %>%
  rowwise() %>%
  mutate(
    openalex_id = stringr::str_extract(id, "W\\d+"),
    institutions = list(unique(author$institution_display_name)),
    .keep = "none"
  )

multi_insts

# institutions of the first work
str(multi_insts[1, "institutions"])
```

### List output

If we want to get *all* the institutions that the authors of these works are affiliated with (since one author may be affiliated with more than one institution), you would want to work with the list output.

As you can see, the nested list can be more convoluted to work with:

```{r}
works_list <- oa_fetch(
  entity = "works",
  title.search = c("bibliometric analysis", "science mapping"),
  cited_by_count = ">100",
  from_publication_date = "2020-01-01",
  to_publication_date = "2021-01-31",
  sort = "cited_by_count:desc",
  output = "list"
)

work_authors <- lapply(works_list, \(x) x$author)

unique_insts <- sapply(
  work_authors,
  \(z) unique(unlist(
    sapply(
      z, \(y) sapply(y$institutions, \(x) x$display_name)
    )
  ))
)

unique_insts[[1]]
```
