---
title: "Filters"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


In your `oa_fetch` calls, you can specify additional arguments to *filter* your search result.
Available arguments/filters for each entity and associated example values are in the tables below.

**Note**: `x_concepts.id` in all entities are experimental.

See the [original documentation](https://docs.openalex.org/how-to-use-the-api/get-lists-of-entities/filter-entity-lists) from Open Alex.

## Available arguments by entity

### Works

| Filter                                  | Example value                                                                              |
|-----------------------------------------|--------------------------------------------------------------------------------------------|
| `publication_year`                      | 2018                                                                                       |
| `publication_date`                      | \"2018-02-13\"                                                                             |
| `primary_location.source.issn`          | \"2167-8359\"                                                                              |
| `primary_location.license`              | \"cc-by\"                                                                                  |
| `primary_location.source.host_organization` | \"<https://openalex.org/P4310320104>\"                                                                                  |
| `primary_location.source.type`          | \"journal\"                                                                                |
| `type`                                  | \"journal-article\"                                                                        |
| `is_paratext`                           | TRUE                                                                                       |
| `open_access.oa_status`                 | \"gold\"                                                                                   |
| `open_access.is_oa`                     | TRUE                                                                                       |
| `authorships.author.id`                 | \"A1969205032\"                                                                            |
| `authorships.author.orcid`              | \"0000-0003-1613-5981\"                                                                    |
| `authorships.institutions.id`           | \"I4200000001\"                                                                            |
| `authorships.institutions.ror`          | \"02nr0ka47\"                                                                              |
| `authorships.institutions.country_code` | \"US\"                                                                                     |
| `authorships.institutions.type`         | \"nonprofit\"                                                                              |
| `cited_by_count`                        | 382                                                                                        |
| `is_retracted`                          | FALSE                                                                                      |
| `concepts.id`                           | \"C2778793908\"                                                                            |
| `concepts.wikidata`                     | \"<https://www.wikidata.org/wiki/Q5122404>\"                                               |
| `doi`                                   | \"10.7717/peerj.4375\"                                                                     |
| `ids.mag`                               | 2741809807                                                                                 |
| `ids.pmid`                              | \"<https://pubmed.ncbi.nlm.nih.gov/23638343/>\"                                            |
| `ids.pmcid`                             | \"<https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3628373/>\"**                              |
| `alternate_host_venues.license`         | \"cc-by\"                                                                                  |
| `alternate_host_venues.version`         | \"submittedVersion\"                                                                       |
| `alternate_host_venues.venue_id`        | \"V1983995261\"                                                                            |
| `display_name.search`                   | \"wombat\"                                                                                 |
| `abstract.search`                       | \"artificial intelligence\"                                                                |
| `fulltext.search`                       | \"climate change\"                                                                         |
| `raw_affiliation_string.search`         | \"department of political science university of Amsterdam\"                                |
| `has_abstract`                          | TRUE                                                                                       |
| `has_doi`                               | FALSE                                                                                      |
| `has_ngrams`                            | TRUE                                                                                       |
| `has_references`                        | TRUE                                                                                       |
| `cites`                                 | \"W2741809807\"                                                                            |
| `cited_by`                              | \"W2766808518\"                                                                            |
| `related_to`                            | \"W2486144666\"                                                                            |
| `from_publication_date`                 | \"2022-08-01\"                                                                             |
| `to_publication_date`                   | \"2022-08-11\"                                                                             |
| `has_oa_accepted_or_published_version`  | TRUE                                                                                       |
| `has_oa_submitted_version`              | TRUE                                                                                       |


### Authors

| Filter                                  | Example value                                                                              |
|-----------------------------------------|--------------------------------------------------------------------------------------------|
| `works_count`                           | \">99\"                                                                                    |
| `cited_by_count`                        | \">1000\"                                                                                  |
| `last_known_institutions.id`            | \"I4200000001\"                                                                            |
| `last_known_institutions.ror`           | \"02nr0ka47\"                                                                              |
| `last_known_institutions.country_code`  | \"CA\"                                                                                     |
| `last_known_institutions.type`          | \"nonprofit\"                                                                              |
| `openalex`                              | \"A2208157607\"                                                                            |
| `orcid`                                 | \"0000-0001-6187-6610\"                                                                    |
| `mag`                                   | \"2208157607\"                                                                             |
| `twitter`                               | \"jasonpriem\"                                                                             |
| `wikipedia`                             | \"<https://en.wikipedia.org/wiki/Jennifer_Doudna>\"                                        |
| `scopus`                                | \"<http://www.scopus.com/inward/authorDetails.url?authorID=36455008000&partnerID=MN8TOARS>\" |
| `x_concepts.id`                         | \"C41008148\"                                                                              |
| `display_name.search`                   | \"tupolev\"                                                                                |
| `search`                                | \"Phillip Kuo\"                                                                            |
| `has_orcid`                             | TRUE                                                                                       |

### Sources

| Filter                                  | Example value                                                                              |
|-----------------------------------------|--------------------------------------------------------------------------------------------|
| `issn`                                  | \"2167-8359\"                                                                              |
| `publisher`                             | \"Peerj\"                                                                                  |
| `works_count`                           | 20184                                                                                      |
| `cited_by_count`                        | 133702                                                                                     |
| `x_concepts.id`                         | \"C185592680\"                                                                             |
| `is_oa`                                 | TRUE                                                                                       |
| `is_in_doaj`                            | TRUE                                                                                       |
| `openalex`                              | \"V1983995261\"                                                                            |
| `issn_l`                                | \"2167-8359\"                                                                              |
| `issn`                                  | \"2167-8359\"                                                                              |
| `mag`                                   | 1983995261                                                                                 |
| `display_name.search`                   | \"Neurology\"                                                                              |
| `has_issn`                              | FALSE                                                                                      |

### Institutions

| Filter                                  | Example value                                                                              |
|-----------------------------------------|--------------------------------------------------------------------------------------------|
| `country_code`                          | \"CN\"                                                                                     |
| `type`                                  | \"education\"                                                                              |
| `works_count`                           | \"<999\"                                                                                   |
| `cited_by_count`                        | \">20000\"                                                                                 |
| `x_concepts.id`                         | \"C41008148\"                                                                              |
| `display_name.search`                   | \"technology\"                                                                             |
| `has_ror`                               | FALSE                                                                                      |
| `openalex`                              | \"C41008148\"                                                                              |
| `ror`                                   | \"0130frc33\"                                                                              |
| `mag`                                   | \"114027177\"                                                                              |
| `grid`                                  | \"grid.10698.36\"                                                                          |
| `wikipedia`                             | \"<https://en.wikipedia.org/wiki/University%20of%20North%20Carolina%20at%20Chapel%20Hill>\"|
| `wikidata`                              | \"<https://www.wikidata.org/wiki/Q192334>\"                                                |

### Concepts

| Filter                                  | Example value                                                                              |
|-----------------------------------------|--------------------------------------------------------------------------------------------|
| `level`                                 | 3                                                                                          |
| `works_count`                           | \"<999\"                                                                                   |
| `cited_by_count`                        | \">10000\"                                                                                 |
| `ancestors.id`                          | \"C161191863\"                                                                             |
| `openalex`                              | \"C2522767166\"                                                                            |
| `wikidata_id`                           | \"Q14565201\"                                                                              |
| `mag`                                   | \"2778407487114027177\"                                                                    |
| `wikipedia`                             | \"<https://en.wikipedia.org/wiki/Altmetrics>\"                                             |
| `umls_aui`                              |                                                                                            |
| `umls_cui`                              |                                                                                            |
| `display_name.search`                   | \"electrodynamics\"                                                                        |
| `has_wikidata`                          | FALSE                                                                                      |



## Examples

```{r examples}
library(openalexR)

# Unlike the other filters, search does NOT require an exact match. 
# This is particularly useful to search for authors. Some authors have their middle names in a variety of forms, which may not exist, or co-exist in OpenAlex (e.g. Phillip H. Kuo, Phillip Hsin Kuo).  
# The display_name search returns an exact match, and will NOT find all these variations. For example, author "Phillip H. Kuo" and "Phillip Hsin Kuo" is the same person.
# His middle name are recorded differently in openAlex. Therefore, all the varitions can only be found either using search ="Phillip Kuo" or display_name =c("Phillip H. Kuo" , "Phillip Hsin Kuo").

authors_from_names <- oa_fetch(
  entity = "authors", 
  search = "Phillip Kuo"
 )


lib_concepts <- oa_fetch(
  entity = "concepts", 
  level = 3,
  works_count = "<999",
  cited_by_count = ">10000",
  ancestors.id = "C161191863",
  has_wikidata = TRUE
)
dplyr::glimpse(lib_concepts)

tech_insts <- oa_fetch(
  entity = "institutions", 
  country_code = "CN",
  x_concepts.id = "C41008148",
  display_name.search = "technology",
  has_ror = FALSE
)
dplyr::glimpse(tech_insts)

peer_venues <- oa_fetch(
  entity = "sources", 
  display_name = "PeerJ",
  x_concepts.id = "C185592680",
  is_oa = TRUE,
  is_in_doaj = TRUE
)
dplyr::glimpse(peer_venues)
```

