---
title: "Transition towards Open Access"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dpi = 300
)
```

Following the template in OpenAlex's [oa-percentage tutorial](https://github.com/ourresearch/openalex-api-tutorials/blob/develop/notebooks/institutions/oa-percentage.ipynb), this vignette uses openalexR to answer:

**How many of recent journal articles from the University of Pennsylvania are open access? And how many aren't?**

```{r setup, message=FALSE}
library(openalexR)
library(dplyr)
library(tidyr)
library(ggplot2)
```

We first need to find the [openalex.id](https://docs.openalex.org/how-to-use-the-api/get-single-entities#the-openalex-id) for University of Pennsylvania.
We can do this by fetching for the *institutions* `entity` and put *"University of Pennsylvania"* in `display_name` or `display_name.search`:

```{r}
oa_fetch(
  entity = "inst", # same as "institutions"
  display_name.search = "\"University of Pennsylvania\""
) %>%
  select(display_name, ror) %>% 
  knitr::kable()
```

We will use the first ror, [00b30xv10](https://ror.org/00b30xv10), as one of the filters for our query.

***Alternatively***, we could go to the autocomplete endpoint at <https://explore.openalex.org/> to search for "University of Pennsylvania" and find the ror [there](https://explore.openalex.org/institutions/I79576946)!

All other filters are straightforward and explained in detailed in the original jupyter notebook [tutorial](https://github.com/ourresearch/openalex-api-tutorials/blob/develop/notebooks/institutions/oa-percentage.ipynb).
The only difference here is that, instead of grouping by `is_oa`, we're interested in the "trend" over the years, so we're going to group by `publication_year`, and perform the query twice, one for `is_oa = "true"` and one for `is_oa = "false"` .

```{r}
open_access <- oa_fetch(
  entity = "works",
  institutions.ror = "00b30xv10",
  type = "article",
  from_publication_date = "2012-08-24",
  is_paratext = "false",
  is_oa = "true",
  group_by = "publication_year"
)

closed_access <- oa_fetch(
  entity = "works",
  institutions.ror = "00b30xv10",
  type = "article",
  from_publication_date = "2012-08-24",
  is_paratext = "false",
  is_oa = "false",
  group_by = "publication_year"
)

uf_df <- closed_access %>%
  select(- key_display_name) %>%
  full_join(open_access, by = "key", suffix = c("_ca", "_oa")) 

uf_df
```

Finally, we compare the number of open vs. closed access articles over the years:

```{r warning=FALSE}
uf_df %>%
  filter(key <= 2021) %>% # we do not yet have complete data for 2022 and after
  pivot_longer(cols = starts_with("count")) %>%
  mutate(
    year = as.integer(key),
    is_oa = recode(
      name,
      "count_ca" = "Closed Access",
      "count_oa" = "Open Access"
    ),
    label = if_else(key < 2021, NA_character_, is_oa)
  ) %>% 
  select(year, value, is_oa, label) %>%
  ggplot(aes(x = year, y = value, group = is_oa, color = is_oa)) +
  geom_line(size = 1) +
  labs(
    title = "University of Pennsylvania's progress towards Open Access",
    x = NULL, y = "Number of journal articles") +
  scale_color_brewer(palette = "Dark2", direction = -1) +
  scale_x_continuous(breaks = seq(2010, 2024, 2)) +
  geom_text(aes(label = label), nudge_x = 0.1, hjust = 0) +
  coord_cartesian(xlim = c(NA, 2022.5)) +
  guides(color = "none")
```
